# Modeling Diabetes Occurrence

```{r message = FALSE, warning= FALSE}
library(dplyr)
library(tidyverse)
library(tibble)
library(ggplot2)
library(data.table)
library(DT)
library(broom)
library(devtools)
# library(viridis)
library(knitr)
# library(hrbrthemes)
library(car)
#using data specified in this github repository:
install_github("jhs-hwg/cardioStatsUSA")
library(cardioStatsUSA)
```

## Simple Diabetes Modeling

We are choosing Diabetes as the co-morbidity for modeling. We will try to predict the occurrence of diabetes as a function of blood pressure as well as other covariates.

The most basic model possible is a simple logistic regression model predicting diabetes status based on blood pressure:

```{r}
#to prevent errors, exclude the rows with na:
used_vars = c('cc_diabetes', 'bp_sys_mean', 'demo_age_years', 'demo_race', 'demo_gender', 'cc_bmi', 'cc_smoke', 'bp_med_use')
clean_nhanes <- nhanes_data[complete.cases(nhanes_data[,..used_vars]), ]

simple_diab_model_sys <- glm(cc_diabetes ~ bp_sys_mean, data = clean_nhanes, family = binomial)

summary(simple_diab_model_sys)
knitr::kable(tidy(simple_diab_model_sys, exponentiate = TRUE, conf.int = TRUE), digits = 3)

simple_diab_model_dia <- glm(cc_diabetes ~ bp_dia_mean, data = clean_nhanes, family = binomial)

summary(simple_diab_model_dia)
knitr::kable(tidy(simple_diab_model_dia, exponentiate = TRUE, conf.int = TRUE), digits = 3)
```

#### Model Explanation:

For the logistic regression using systolic blood pressure, we can see that the odds ratio beta parameter is 1.024. This implies that a higher systolic blood pressure corresponds to increased chances of observing diabetes.

A more exact interpretation would be that: for one additional unit increase in systolic blood pressure, the estimated risk of having diabetes increases by 2.4%.

This value may not seem extremely high, yet it still has a p value evaluated to be 0 by R. We can also see that the 95% confidence interval of this parameter ranges between 1.022 and 1.024, meaning that the model predicts that the true relationship falls within this interval with a 95% certainty, since the true value of the parameter is most likely within this interval.

The opposite is true with diastolic blood pressure: the beta parameter is 0.996, implying that higher diastolic blood pressure corresponds to a lower chance of observing diabetes. For an additional unit increase in diastolic blood pressure, the estimated risk of having diabetes decreases by 0.04%.

Note: systolic blood pressure is a better predictor of cardiovascular disease, so I chose to consider systolic rather than diastolic in future models.

## Demographic Model

Next, we build a model including the demographic variables as well. This will hopefully provide a more accurate model, since it will have access to the further information in order to make predictions.

```{r}
demo_diab_model_sys <- glm(cc_diabetes ~ bp_sys_mean + demo_age_years + demo_race + demo_gender,data = clean_nhanes, family = binomial)

summary (demo_diab_model_sys)
knitr::kable(tidy(demo_diab_model_sys, exponentiate = TRUE, conf.int = TRUE), digits = 3)
```

#### Model Explanation

In this model, we can still see a slight upwards trend with systolic blood pressure, although it is a bit weaker (1.024 \> 1.003). With this parameter, the estimated risk of having diabetes increases only by 0.3%. Although the effect is smaller, we can see that the confidence interval is between 1.003 and 1.005, which does not include 1. So, although small, the positive relationship is most likely significant.

#### Demographics

**Age:**

The model also tells us that age is positively correlated with the observation of diabetes. With a log odds ratio of 1.054, it seems that every year increases the risk of diabetes occurrence by 5.4%.

**Race:**

The log odds ratio for categorical variables compares the other variables to the baseline "reference group". Here, the "reference group" for the race variable is Non-Hispanic White. We can see that, in comparison to this group, all other races have significantly higher risk of diabetes: for example, the Hispanic population's risk of diabetes occurrence is 2.36 times higher than the Non-Hispanic White population.

**Gender:**

With this variable, the reference group is men. We can see that women have a lower odds of having diabetes compared to men. Specifically, the risk of diabetes in women is about 85.6% of that of men.

## Full Model

Lastly, we want to create a model including confounding variables. To achieve this, we must first determine which other variables are confounding, meaning they have significant associations with diabetes.

We can achieve this by using chi squared test, which compares whether or not the observed ratios of diabetes and confounding variables matches the ratios expected by random chance. If the chi squared test determines significance, then diabetes and the other variables commonly occur together and therefore are confounding.

#### Perform Chi Squared Tests:

```{r}
#Testing for BMI
nhanes_bmi_diab <- nhanes_data %>% select(cc_diabetes, cc_bmi) %>% drop_na(cc_bmi) %>% drop_na(cc_diabetes)
nhanes_bmi_diab <- table(nhanes_bmi_diab)
#nhanes_bmi_diab
chisq.test(nhanes_bmi_diab)

#Testing for Smoking
nhanes_smoke_diab <- nhanes_data %>% select(cc_diabetes, cc_smoke) %>% drop_na(cc_smoke) %>% drop_na(cc_diabetes)
nhanes_smoke_diab <- table(nhanes_smoke_diab)
#nhanes_smoke_diab
chisq.test(nhanes_smoke_diab)

#Testing for Hypertensive Medication use
nhanes_med_diab <- nhanes_data %>% select(cc_diabetes, bp_med_use) %>% drop_na(bp_med_use) %>% drop_na(cc_diabetes)
nhanes_med_diab <- table(nhanes_med_diab)
#nhanes_med_diab
chisq.test(nhanes_med_diab)
```

After this analysis, we can see that all three variables, BMI, Smoking, and Medication Use, are all confounding. We can include all of them in the next model, in addition to systolic blood pressure and the demographic variables:

```{r}
full_diab_model_sys <- glm(cc_diabetes ~ bp_sys_mean + demo_age_years + demo_race + demo_gender+ cc_bmi + cc_smoke + bp_med_use, data = clean_nhanes, family = binomial)

summary(full_diab_model_sys)
knitr::kable(tidy(full_diab_model_sys, exponentiate = TRUE, conf.int = TRUE), digits = 3)
```

#### Model Explanation

The previous variables in this model continue to show the same patterns, although again, the effect of systolic blood pressure is again reduced due to a lower log odds ratio. Now, the p value seems to be much higher at 0.02 (which is still significant, but much less so). The demographic variables also have much the same pattern, although Non-Hispanic Asians seem to have a much higher odds ratio than in the previous model.

#### Covariates

**BMI**:

In this variable, the baseline reference level is bmi \<25. In comparison, those with higher bmi have increased risk of having diabetes, with the risk increasing the higher the bmi becomes. For example, those with bmi from 25-30 have a 1.8 times higher risk, while those with a bmi of 35+ have a 5.8 times higher risk.

**Smoking:**

With a baseline level of nonsmoking, former smokers may have about a 6% higher risk of diabetes. This may not be significant, however, given that the p value is only 0.067 and the odds ratio of 1 falls within the confidence interval. Current smokers, on the other hand, have about a 20% higher risk, this time with a much lower p value of 4\*10\^-6

**Blood Pressure Medication Use**:

Compared to those who do not use blood pressure medication, individuals who do have about a 2.5 times higher risk of experiencing diabetes.

## Model Selection

It is possible to find an even better model by modifying which variables we include in our logistic regression model. We can achieve this using model selection, which will selectively add or subtract our explanatory variables from the model and evaluate its performance to choose the final best set of variables.

Explanation of model selection and or AIC?

#### Backwards Selection:

```{r}
backwards = step(full_diab_model_sys)
formula(backwards)
```

It looks like the backwards selection model returned the same full model. In other words, removing variables does not improve its AIC, and therefore does not improve its performance.

#### Forward Selection

```{r}
#start from a model without any predictors
nothing <- glm(cc_diabetes ~ 1, data = clean_nhanes, family=binomial)
summary(nothing)

forwards = step(nothing,
scope=list(lower=formula(nothing),upper=formula(full_diab_model_sys)), direction="forward")
```

The Forward Selection process again selects the full model.

### Forward-Backward Selection

```{r}
bothways = step(nothing, list(lower=formula(nothing),upper=formula(full_diab_model_sys)),
direction="both",trace=0)
formula(bothways)
```

The Forward-Backward algorithm also selects all of the variables.

In this case, the best model (as determined by the AIC) includes all of the variables (systolic blood pressure, age, race, sex, smoking status, BMI, and medication use). In other situations, it may have returned just a few of these variables, but for this example, we can proceed with the full model.

# Modeling for Subset of Data

```{r}
#Read in the subset data
subset_nhanes <- readRDS("nhanes_subset.rds")
```

## Simple Diabetes Modeling

We are choosing Diabetes as the co-morbidity for modeling. We will try to predict the occurrence of diabetes as a function of blood pressure as well as other covariates.

The most basic model possible is a simple logistic regression model predicting diabetes status based on blood pressure:

```{r}
#to prevent errors, exclude the rows with na:
used_vars = c('cc_diabetes', 'bp_sys_mean', 'demo_age_years', 'demo_race', 'demo_gender', 'cc_bmi', 'cc_smoke', 'bp_med_use')
clean_subset <- subset_nhanes[,..used_vars]
clean_subset

simple_diab_model_subset <- glm(cc_diabetes ~ bp_sys_mean, data = clean_subset, family = binomial)

summary(simple_diab_model_subset)
knitr::kable(tidy(simple_diab_model_subset, exponentiate = TRUE, conf.int = TRUE), digits = 3)

#simple_diab_model_dia <- glm(cc_diabetes ~ bp_dia_mean, data = clean_nhanes, family = binomial)

#summary(simple_diab_model_dia)
#knitr::kable(tidy(simple_diab_model_dia, exponentiate = TRUE, conf.int = TRUE), digits = 3)
```

#### Model Explanation:

For the logistic regression using systolic blood pressure, we can see that the odds ratio beta parameter is 1.024. This implies that a higher systolic blood pressure corresponds to increased chances of observing diabetes.

A more exact interpretation would be that: for one additional unit increase in systolic blood pressure, the estimated risk of having diabetes increases by 2.4%.

This value may not seem extremely high, yet it still has a p value evaluated to be 0 by R. We can also see that the 95% confidence interval of this parameter ranges between 1.022 and 1.024, meaning that the model predicts that the true relationship falls within this interval with a 95% certainty, since the true value of the parameter is most likely within this interval.

The opposite is true with diastolic blood pressure: the beta parameter is 0.996, implying that higher diastolic blood pressure corresponds to a lower chance of observing diabetes. For an additional unit increase in diastolic blood pressure, the estimated risk of having diabetes decreases by 0.04%.

Note: systolic blood pressure is a better predictor of cardiovascular disease, so I chose to consider systolic rather than diastolic in future models.

## Demographic Model

Next, we build a model including the demographic variables as well. This will hopefully provide a more accurate model, since it will have access to the further information in order to make predictions.

```{r}
demo_diab_model_subset <- glm(cc_diabetes ~ bp_sys_mean + demo_age_years + demo_race + demo_gender,data = clean_subset, family = binomial)

summary (demo_diab_model_subset)
knitr::kable(tidy(demo_diab_model_subset, exponentiate = TRUE, conf.int = TRUE), digits = 3)
```

#### Model Explanation

In this model, we can still see a slight upwards trend with systolic blood pressure, although it is a bit weaker (1.024 \> 1.003). With this parameter, the estimated risk of having diabetes increases only by 0.3%. Although the effect is smaller, we can see that the confidence interval is between 1.003 and 1.005, which does not include 1. So, although small, the positive relationship is most likely significant.

#### Demographics

**Age:**

The model also tells us that age is positively correlated with the observation of diabetes. With a log odds ratio of 1.054, it seems that every year increases the risk of diabetes occurrence by 5.4%.

**Race:**

The log odds ratio for categorical variables compares the other variables to the baseline "reference group". Here, the "reference group" for the race variable is Non-Hispanic White. We can see that, in comparison to this group, all other races have significantly higher risk of diabetes: for example, the Hispanic population's risk of diabetes occurrence is 2.36 times higher than the Non-Hispanic White population.

**Gender:**

With this variable, the reference group is men. We can see that women have a lower odds of having diabetes compared to men. Specifically, the risk of diabetes in women is about 85.6% of that of men.

## Full Model

Lastly, we want to create a model including confounding variables. To achieve this, we must first determine which other variables are confounding, meaning they have significant associations with diabetes.

We can achieve this by using chi squared test, which compares whether or not the observed ratios of diabetes and confounding variables matches the ratios expected by random chance. If the chi squared test determines significance, then diabetes and the other variables commonly occur together and therefore are confounding.

#### Perform Chi Squared Tests:

```{r}
#Testing for BMI
nhanes_bmi_diab_subset <- clean_subset %>% select(cc_diabetes, cc_bmi) %>% drop_na(cc_bmi) %>% drop_na(cc_diabetes)
nhanes_bmi_diab_subset <- table(nhanes_bmi_diab_subset)
#nhanes_bmi_diab
chisq.test(nhanes_bmi_diab_subset)

#Testing for Smoking
nhanes_smoke_diab_subset <- clean_subset %>% select(cc_diabetes, cc_smoke) %>% drop_na(cc_smoke) %>% drop_na(cc_diabetes)
nhanes_smoke_diab_subset <- table(nhanes_smoke_diab_subset)
#nhanes_smoke_diab
chisq.test(nhanes_smoke_diab_subset)

#Testing for Hypertensive Medication use
nhanes_med_diab_subset <- clean_subset %>% select(cc_diabetes, bp_med_use) %>% drop_na(bp_med_use) %>% drop_na(cc_diabetes)
nhanes_med_diab_subset <- table(nhanes_med_diab_subset)
#nhanes_med_diab
chisq.test(nhanes_med_diab_subset)
```

After this analysis, we can see that all three variables, BMI, Smoking, and Medication Use, are all confounding. We can include all of them in the next model, in addition to systolic blood pressure and the demographic variables:

```{r}
full_diab_model_subset <- glm(cc_diabetes ~ bp_sys_mean + demo_age_years + demo_race + demo_gender+ cc_bmi + cc_smoke + bp_med_use, data = clean_subset, family = binomial)

summary(full_diab_model_subset)
knitr::kable(tidy(full_diab_model_subset, exponentiate = TRUE, conf.int = TRUE), digits = 3)
```

#### Model Explanation

The previous variables in this model continue to show the same patterns, although again, the effect of systolic blood pressure is again reduced due to a lower log odds ratio. Now, the p value seems to be much higher at 0.02 (which is still significant, but much less so). The demographic variables also have much the same pattern, although Non-Hispanic Asians seem to have a much higher odds ratio than in the previous model.

#### Covariates

**BMI**:

In this variable, the baseline reference level is bmi \<25. In comparison, those with higher bmi have increased risk of having diabetes, with the risk increasing the higher the bmi becomes. For example, those with bmi from 25-30 have a 1.8 times higher risk, while those with a bmi of 35+ have a 5.8 times higher risk.

**Smoking:**

With a baseline level of nonsmoking, former smokers may have about a 6% higher risk of diabetes. This may not be significant, however, given that the p value is only 0.067 and the odds ratio of 1 falls within the confidence interval. Current smokers, on the other hand, have about a 20% higher risk, this time with a much lower p value of 4\*10\^-6

**Blood Pressure Medication Use**:

Compared to those who do not use blood pressure medication, individuals who do have about a 2.5 times higher risk of experiencing diabetes.

## Model Selection

It is possible to find an even better model by modifying which variables we include in our logistic regression model. We can achieve this using model selection, which will selectively add or subtract our explanatory variables from the model and evaluate its performance to choose the final best set of variables.

Explanation of model selection and or AIC?

#### Backwards Selection:

```{r}
backwards = step(full_diab_model_subset)
formula(backwards)
```

It looks like the backwards selection model returned the same full model. In other words, removing variables does not improve its AIC, and therefore does not improve its performance.

#### Forward Selection

```{r}
#start from a model without any predictors
nothing <- glm(cc_diabetes ~ 1, data = subset_nhanes, family=binomial)
summary(nothing)

forwards = step(nothing,
scope=list(lower=formula(nothing),upper=formula(full_diab_model_subset)), direction="forward")
```

The Forward Selection process again selects the full model.

### Forward-Backward Selection

```{r}
bothways = step(nothing, list(lower=formula(nothing),upper=formula(full_diab_model_subset)),
direction="both",trace=0)
formula(bothways)
```

The Forward-Backward algorithm also selects all of the variables.

In this case, the best model (as determined by the AIC) includes all of the variables (systolic blood pressure, age, race, sex, smoking status, BMI, and medication use). In other situations, it may have returned just a few of these variables, but for this example, we can proceed with the full model.

## 
