# Regression Calibration

```{r message = FALSE, warning= FALSE}
library(dplyr)
library(tidyverse)
library(tibble)
library(ggplot2)
library(knitr)
library(hrbrthemes)
library(viridis)

library(devtools)
#using data specified in this github repository:
install_github("jhs-hwg/cardioStatsUSA")
library(cardioStatsUSA)
```

```{r}
#Read in the subset data
subset_nhanes <- readRDS("nhanes_subset.rds")
```

## Naive Analysis

Before starting with regression calibration, we first want to see what a model that only uses our noisy data looks like. We'll use the same model formula as before (including age, race, smoking status etc.), the only difference being that we will use our error-full data, specifically the low reliability case.

```{r}
naive.model=glm(cc_diabetes ~ bp_sys_mean_noise_low + demo_age_years + demo_race + demo_gender+ cc_bmi + cc_smoke + bp_med_use, family="binomial", data=subset_nhanes)
summary(naive.model)
knitr::kable(tidy(naive.model, exponentiate = TRUE, conf.int = TRUE), digits = 3)
```

Looking at the summary above, the significant coefficients, and therefore the variables which the model finds most useful in diabetes status prediction, are age, Non-Hispanic Black racial status, and BMI status, specifically for those in the 25 to 30 range. Note that our p value for our blood pressure variable is quite high, showing non-significance in the model.

## Gold Standard Analysis

Next, we'll look at our "true" model, the model which uses our unbiased, non-error blood pressure values:

```{r}
true.model=glm(cc_diabetes ~ bp_sys_mean + demo_age_years + demo_race + demo_gender+ cc_bmi + cc_smoke + bp_med_use, family="binomial", data=subset_nhanes)
summary(true.model)
knitr::kable(tidy(true.model, exponentiate = TRUE, conf.int = TRUE), digits = 3)
```

For this subset of the data, the true beta value for the systolic mean is ... This is much higher than the model which we looked at previously, using all \~50,000 subjects, although it is to be expected since we restricted the data to n=150 samples. Also notice that the significance level

## Regression Calibration

```{r}
#Create a model with predicts the true BP value based on the noisy BP data
rcfit<-lm(bp_sys_mean ~ bp_sys_mean_noise_low, data=subset_nhanes)
#Add in to lm(): + demo_age_years + demo_race + demo_gender+ cc_bmi + cc_smoke + bp_med_use
summary(rcfit)
subset_nhanes$bp_hat <- predict(rcfit, newdata=subset_nhanes) 

final.model = glm(cc_diabetes ~ bp_hat + demo_age_years + demo_race + demo_gender+ cc_bmi + cc_smoke + bp_med_use, family="binomial",data=subset_nhanes)
summary(final.model)
knitr::kable(tidy(final.model, exponentiate = TRUE, conf.int = TRUE), digits = 3)

```

```{r}
print(paste("Correlation between BP_noise and BP_hat:", cor(subset_nhanes$bp_sys_mean_noise_low, subset_nhanes$bp_hat)))
```
