# Download NHANES Data

Download the NHANES data using the nhanesA package:

```{r include = FALSE}
library(DT)
library(dplyr)
library(nhanesA)
```

## Identify Required Data Tables:

Look at the diet, examination, and demographic info for the 2011-2012 cycle of NHANES

```{r}
#Note DR1TOT_G and DR2TOT_G
datatable(nhanesTables('DIET', 2012))
#Note BPX_G: Blood Pressure
datatable(nhanesTables('EXAM', 2012))
#Note DEMO_G: Demographic Info (the only table)
datatable(nhanesTables('DEMO', 2012))
```

Now that we have identified the tables we need, take a look at the required variables

```{r}
datatable(nhanesTableVars('DIET', 'DR1TOT_G'))
```

We have identified the variables corresponding to each variable of interest:

| Dietary Measure | Variable Name |
|-----------------|---------------|
| Sodium (mg)     | DR1TSODI      |
| Potassium (mg)  | DR1TPOTA      |
| Calcium (mg)    | DR1TCALC      |
| Magnesium (mg)  | DR1TMAGN      |
| Protein (gm)    | DR1TPROT      |
| Alcohol (gm)    | DR1TALCO      |
| Energy (kcal)   | DR1TKCAL      |

Also include SEQN, which corresponds to to the respondent ID number.

```{r}
datatable(nhanesTableVars('EXAM', 'BPX_G'))
```

We are interested in BPXSY1 and BPXDI1 for systolic and diastolic blood pressure, respectively

## Merge Data Together

Take the diet data

```{r}
d1_diet_d <- nhanes('DR1TOT_G')
d2_diet_d <- nhanes('DR2TOT_G')

#concatenate the diet data for two days together
diet_d <- merge(d1_diet_d, d2_diet_d, )
datatable(head(diet_d))
```

Select Day 1 diet measures

```{r}
select_cols <- c('SEQN', 'DR1TSODI', 'DR1TPOTA', 'DR1TCALC', 'DR1TMAGN', 'DR1TPROT', 'DR1TALCO', 'DR1TKCAL')
nutrients <- diet_d[, select_cols]
datatable(head(nutrients))
```

```{r}
demo <- nhanes('DEMO_G')
head(demo)
```

Select Blood Pressure Measures

```{r}
BP_cols <- c('SEQN', 'BPXSY1','BPXDI1')
BP_data <- nhanes('BPX_G')[, BP_cols]
datatable(head(BP_data))
```

Combine all tables together:

```{r}
total_df <- merge(nutrients, BP_data, by = 'SEQN')
#get rid of any column with NA
total_df <- na.omit(total_df)

total_df <- merge(total_df, demo)
datatable(head(total_df))
```

```{r}
all_v_diastolic_model = lm(formula = BPXDI1 ~ DR1TSODI + DR1TPOTA + DR1TCALC + DR1TMAGN + DR1TPROT + DR1TALCO + DR1TKCAL, data = total_df)

summary(all_v_diastolic_model)
```
