# Preparing the NHANES Data

Download the NHANES data using the nhanesA package:

```{r include = FALSE}
library(DT)
library(dplyr)
library(nhanesA)
```

## Identify Data Tables of Interest

Look at the diet, examination, and demographic info for the 2011-2012 cycle of NHANES

```{r}
#Note DR1TOT_G and DR2TOT_G
datatable(nhanesTables('DIET', 2012))
#Note BPX_G: Blood Pressure (may also need BMX_G/body measures)
datatable(nhanesTables('EXAM', 2012))
#Note DEMO_G: Demographic Info (the only table)
datatable(nhanesTables('DEMO', 2012))
```

## Identify Variables of Interest

Now that we have identified the tables we need, first take a look at the required variables in the table of dietary information:

```{r}
datatable(nhanesTableVars('DIET', 'DR1TOT_G'))
```

We can then identify the variables corresponding to each dietary measure of interest:

| Dietary Measure | Variable Name |
|-----------------|---------------|
| Sodium (mg)     | DR1TSODI      |
| Potassium (mg)  | DR1TPOTA      |
| Calcium (mg)    | DR1TCALC      |
| Magnesium (mg)  | DR1TMAGN      |
| Protein (gm)    | DR1TPROT      |
| Alcohol (gm)    | DR1TALCO      |
| Energy (kcal)   | DR1TKCAL      |

We will also need to include SEQN, which corresponds to to the respondent ID number. This will be the column used for combining the different tables together.

Let's next look at the variables in the blood pressure table:

```{r}
datatable(nhanesTableVars('EXAM', 'BPX_G'))
```

We are interested in BPXSY1 and BPXDI1 for systolic and diastolic blood pressure, respectively

## Merge Data Together

Take the diet data: there are two separate tables for measurements taken on different days. We can combine this using the merge() function.

```{r}
d1_diet_d <- nhanes('DR1TOT_G')
d2_diet_d <- nhanes('DR2TOT_G')

#concatenate the diet data for two days together
diet_d <- merge(d1_diet_d, d2_diet_d, )
datatable(head(diet_d))
```

For this example, we take only the day 1 diet measures to create the "nutrients" dataframe:

```{r}
select_cols <- c('SEQN', 'DR1TSODI', 'DR1TPOTA', 'DR1TCALC', 'DR1TMAGN', 'DR1TPROT', 'DR1TALCO', 'DR1TKCAL')
nutrients <- diet_d[, select_cols]
datatable(head(nutrients))
```

```{r}
demo <- nhanes('DEMO_G')
datatable(head(demo))
```

Simliarly, we select the systolic and diastolic blood pressure measures from the blood pressure table 'BPX_G':

```{r}
BP_cols <- c('SEQN', 'BPXSY1','BPXDI1')
BP_data <- nhanes('BPX_G')[, BP_cols]
datatable(head(BP_data))
```

Combine all tables together:

```{r}
total_df <- merge(nutrients, BP_data, by = 'SEQN')
#get rid of any column with NA
total_df <- na.omit(total_df)

total_df <- merge(total_df, demo)
datatable(head(total_df))
```

This one uses all the variables (but have to adjust for collinearity)

```{r}
all_v_diastolic_model = lm(formula = BPXDI1 ~ DR1TSODI + DR1TPOTA + DR1TCALC + DR1TMAGN + DR1TPROT + DR1TALCO + DR1TKCAL, data = total_df)

summary(all_v_diastolic_model)
```

Note for later:\
Can use the save(dataframe, "file_name.Rdata") function and then in another qmd file load using load("file_name.Rdata")
